### 2024-07-01 FinalThesis/Chapter_01/Manuscript.tex
Publish PHD

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{lineno}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[flushleft]{threeparttable}
\usepackage[nomarkers, nolists]{endfloat}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{scalerel}
\usepackage{tikz}
\usetikzlibrary{svg.path}

%------------------------------------------------------------------------------
%	Some Styling
%------------------------------------------------------------------------------
% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

% Change how equations are referenced
\renewcommand{\theequation}{Equation \arabic{equation}}%

% To be able to put an ORCID
\definecolor{orcidlogocol}{HTML}{A6CE39}
\tikzset{
  orcidlogo/.pic={
    \fill[orcidlogocol] svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3,0,128,0C198.7,0,256,57.3,256,128z};
    \fill[white] svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
                 svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
                 svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
  }
}
\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
\begin{tikzpicture}[yscale=-1,transform shape]
  \pic{orcidlogo};
\end{tikzpicture}
}{|}}}}

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{A Three-Step Approach for Assessing Landscape Connectivity via Simulated
Dispersal: African Wild Dog Case Study}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1} \orcid{0000-0001-7477-2642} \and
  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538}
}

% Reduce spacing between authors
\makeatletter
\def\and{%
  \end{tabular}%
  \hskip -0.5em \@plus.17fil\relax
  \begin{tabular}[t]{c}}
\makeatother

% Current Date
% \date{\today}

% And here the masterpiece begins
\begin{document}

% Change page numbering
\pagenumbering{gobble}

% Required to be able to cite
\bibliographystyle{apalike}

% Create Titlepage
\maketitle

%------------------------------------------------------------------------------
%	Titlepage: Additional Info
%------------------------------------------------------------------------------
\begin{flushleft}

\vspace{0.5cm}

\textsuperscript{1} Department of Evolutionary Biology and Environmental
Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
Switzerland.

\textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
Private Bag 13, Maun, Botswana.

\textsuperscript{\S} Corresponding author (david.hofmann2@uzh.ch)

\vspace{4cm}

\textbf{Running Title:} Simulating Wild Dog Dispersal Trajectories to Assess
Landscape Connectivity

\vspace{0.5cm}

\textbf{Keywords:} dispersal, simulation, movement, integrated step-selection
function, Kavango-Zambezi Transfrontier Conservation Area, landscape
connectivity, Lycaon pictus

\end{flushleft}

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
% \newpage
% \begin{abstract}
% %------------------------------------------------------------------------------
% % Short Version
% %------------------------------------------------------------------------------
% \textbf{Short Version }
% The ability to disperse is contingent on a sufficient degree of landscape
% connectivity, which is why the identification and preservation of movement
% corridors that promote connectivity has become a task of extraordinary
% importance. Currently, ecologists rely on least-cost analysis or circuit theory
% to investigate connectivity, although both methods make several assumptions that
% are hardly met in reality. To address these issues, simulations from
% individual-based movement models have been proposed, yet a unified framework to
% simulate dispersal and quantify connectivity is lacking.
%
% Here, we propose a simple three-step approach that combines several
% already-existing methods to assess connectivity using simulated dispersal
% trajectories. In step one, we use integrated step-selection functions to
% parametrize a mechanistic movement model rendering dispersal behavior. In step
% two, we apply the parametrized model as an individual-based movement model to
% simulate dispersal trajectories. In step three, we combine simulated
% trajectories into three complementary connectivity maps, each focusing on a
% different aspect of landscape connectivity.
%
% We showcase the application of the proposed approach using empirical data of
% dispersing African wild dogs (\textit{Lycaon pictus}) and assess landscape
% connectivity within the world's largest transboundary conservation area in
% Southern Africa. We thereby shed light into dispersing wild dogs' habitat and
% movement preferences, while also uncovering crucial dispersal corridors. With
% this analysis, we demonstrate that simulations from integrated step-selection
% functions offer a simple, yet powerful alternative to traditional connectivity
% modeling techniques.
% \end{abstract}

\newpage
\begin{abstract}
%------------------------------------------------------------------------------
%  LONG VERSION
%------------------------------------------------------------------------------
\noindent \textbf{Context} \quad Dispersal of individuals contributes to
long-term population persistence, yet requires a sufficient degree of landscape
connectivity. To date, connectivity has mainly been investigated using
least-cost analysis and circuit theory, two methods that make assumptions that
are hardly applicable to dispersal. While these assumptions can be relaxed by
explicitly simulating dispersal trajectories across the landscape, a unified
approach for such simulations is lacking.

\noindent \textbf{Objectives} \quad Here, we propose and apply a simple
three-step approach to simulate dispersal and to assess connectivity using
empirical GPS movement data and a set of habitat covariates.

\noindent \textbf{Methods} \quad In step one of the proposed approach, we use
integrated step-selection functions to fit a mechanistic movement model
describing habitat and movement preferences of dispersing individuals. In step
two, we apply the parameterized model to simulate dispersal across the study
area. In step three, we derive three complementary connectivity maps; a heatmap
highlighting frequently traversed areas, a betweenness map pinpointing dispersal
corridors, and a map of inter-patch connectivity indicating the presence and
intensity of functional links between habitat patches. We demonstrate the
applicability of the proposed three-step approach in a case study in which we
use GPS data collected on dispersing African wild dogs (\textit{Lycaon pictus})
inhabiting northern Botswana.

\noindent \textbf{Results} \quad Using step-selection functions we successfully
parametrized a detailed dispersal model that described dispersing individuals'
habitat and movement preferences, as well as potential interactions among the
two. The model substantially outperformed a model that omitted such interactions
and enabled us to simulate 80,000 dispersal trajectories across the study area.

\noindent \textbf{Conclusion} \quad By explicitly simulating dispersal
trajectories, our approach not only requires fewer unrealistic assumptions about
dispersal, but also permits the calculation of multiple connectivity metrics
that together provide a comprehensive view of landscape connectivity. In our
case study, the three derived connectivity maps revealed several wild dog
dispersal hotspots and corridors across the extent of our study area. Each map
highlighted a different aspect of landscape connectivity, thus emphasizing their
complementary nature. Overall, our case study demonstrates that a
simulation-based approach offers a simple yet powerful alternative to
traditional connectivity modeling techniques. It is therefore useful for a
variety of applications in ecological, evolutionary, and conservation research.

\end{abstract}

%------------------------------------------------------------------------------
%	Main Text
%------------------------------------------------------------------------------
\newpage

% \onehalfspacing
% \tableofcontents
\doublespacing

% Change page numbering
\newpage
\pagenumbering{arabic}

% Create linenumbers
\linenumbers

\section{Introduction}

% Importance of Dispersal & Connectivity
% \subsection{Importance of Connectivity \& Connectivity Models}
Dispersal of individuals is a vital process that allows species to maintain
genetic diversity \citep{Perrin.2000, Frankham.2002, Leigh.2012, Baguette.2013,
LaPoint.2013}, rescue non-viable populations \citep{Brown.1977}, and to colonize
unoccupied habitats \citep{Hanski.1999b, MacArthur.2001}. However, the ability
to disperse depends on a sufficient degree of landscape connectivity
\citep{Fahrig.2003, Clobert.2012}, making the identification and protection of
dispersal corridors that promote connectivity a task of fundamental importance
\citep{Doerr.2011, Rudnick.2012}. Identifying dispersal corridors not only
necessitates a comprehensive understanding of the factors that limit dispersal,
but also an appropriate model to estimate connectivity \citep{Baguette.2013,
Vasudev.2015, Hofmann.2021}. To date, the most commonly used connectivity models
are least-cost path analysis (LCPA; \citealp{Adriaensen.2003}) and circuit
theory (CT; \citealp{McRae.2006, McRae.2008}). Unfortunately, both models rest
on assumptions that appear unsuitable for dispersers, thus calling for the
development of alternative approaches. One promising alternative is to assess
landscape connectivity via simulated dispersal trajectories generated from
individual-based movement models (IBMMs, \citealp{Diniz.2019}). However, IBMMs
require a large number of subjective modeling decisions, thus making
among-system comparisons difficult.

% Issues with Both Methods
% \subsection{Issues with Traditional Connectivity Models}
Traditional connectivity models make assumptions that are rarely met for
dispersers. LCPA, for instance, assumes that individuals move towards a
preconceived endpoint and choose a cost-minimizing route accordingly
\citep{Sawyer.2011, Abrahms.2017}. While this assumption may be justifiable for
migrating animals, it is unlikely to hold for dispersers, as dispersers
typically move across unfamiliar territory towards an unknown endpoint
\citep{Koen.2014, Cozzi.2020}. CT, on the contrary, posits that animals move
according to a random walk, entailing that autocorrelation between subsequent
movements cannot be rendered \citep{Diniz.2019}. For dispersers, however,
autocorrelated movements are regularly observed \citep{Cozzi.2020,
Hofmann.2021}, meaning that dispersal trajectories are usually strongly
directional. An interesting generalization that bridges the continuum between
LCPA and CT has been proposed by \cite{Panzacchi.2016} and enables to
capitalize on the merits of both approaches. Despite these and several other
generalizations of LCPA and CT (e.g. \citealp{Pinto.2009, Landguth.2012,
Panzacchi.2016, Brennan.2020}), some shortcomings remain. Most notably, all of
these methods rely on static permeability or resistance surfaces that can't
reflect the temporal dimension of dispersal. This permits statements about the
expected duration for moving between habitat patches \citep{Martensen.2017,
Diniz.2019}.

% What about IBMMs?
% \subsection{What about IBMMs?}
The shortcomings inherent to LCPA and CT can be overcome by simulating dispersal
using IBMMs and by converting simulated trajectories into meaningful measures of
connectivity \citep{Diniz.2019}. In contrast to LCPA and CT, IBMMs allow to
explicitly simulate how individuals move across and interact with the
encountered landscape \citep{Kanagaraj.2013, Clark.2015, Allen.2016,
Hauenstein.2019, Zeller.2020}, as well as to render potential interactions
between movement behavior and habitat conditions \citep{Avgar.2016}. This shifts
the focus towards a more functional view on connectivity
\citep{Tischendorf.2000}. Furthermore, IBMMs generate movement sequentially,
i.e. they generate a series of steps, so that the temporal dimension of
dispersal becomes explicit and allows modeling autocorrelation between
successive steps \citep{Diniz.2019}. Finally, simulations from IBMMs do not
enforce movement or connections towards preconceived endpoints but allow
individuals to adjust their route  ``on the go'', thereby preventing biases
arising from misplaced endpoints. Despite these advantages, a unifying approach
to simulate dispersal and assess connectivity using IBMMs is lacking.
Considering the large number of subjective decisions entailed by IBMMs, an
approach that streamlines and standardizes the application of dispersal
simulations to assess connectivity will, however, be critical to safeguard
comparability among studies.

% Proposed Solution: Three-Step approach
% \subsection{Proposed Solution: Three-Step Approach}
Here, we propose and exemplify a simple three-step approach for simulating
dispersal and assessing landscape connectivity (\Cref{GraphicalAbstract}). In
step one, we combine GPS movement data of dispersing individuals with habitat
covariates to fit a mechanistic movement model via integrated step-selection
functions (ISSFs, \citealp{Avgar.2016}). We chose to use ISSFs because the
framework not only allows inference on the study species' habitat kernel (i.e.
its habitat preferences), but also its movement kernel (i.e. its movement
preferences/capabilities) and potential interactions among the two
\citep{Avgar.2016, Fieberg.2021}. In step two, we use the parametrized movement
model to simulate dispersal across the study area. Comparable simulations have
already been applied to estimate steady-state utilization distributions of
resident individuals \citep{Potts.2013, Signer.2017} and to model landscape
connectivity, yet disregarding interdependencies between habitat and movement
kernels \citep{Clark.2015, Zeller.2020}. Finally, in step three, we convert the
simulated trajectories into three complementary connectivity maps; (i) a heatmap
revealing frequently traversed areas (e.g. \citealp{Hauenstein.2019,
Zeller.2020}), (ii) a betweenness-map delineating dispersal corridors and
bottlenecks (e.g. \citealp{BastilleRousseau.2018}), (iii) and a map of
inter-patch connectivity, depicting the presence and intensity of functional
links between habitat patches, as well as the average dispersal duration
required to realize those connections (e.g. \citealp{Gustafson.1996,
Kanagaraj.2013}).

@@

% Introduction of the study species
% \subsection{Case Study}
We showcase the application of the proposed approach using GPS movement data
collected on dispersing African wild dogs (\textit{Lycaon pictus}). The African
wild dog is a highly mobile species whose population persistence heavily relies
on the availability of large, natural or semi-natural landscapes and a
sufficient degree of connectivity among remaining subpopulations. Once common
throughout sub-Saharan Africa, this species has disappeared from much of its
historic range, largely due to human persecution, habitat fragmentation, and
disease outbreaks \citep{Woodroffe.2012}. Wild dogs typically disperse in
single-sex coalitions \citep{McNutt.1996, Behr.2020} and are capable of
dispersing several hundred kilometers \citep{DaviesMostert.2012, Masenga.2016,
Cozzi.2020}. Although previous studies have investigated connectivity for this
species using LCPA \citep{Hofmann.2021} and CT \citep{Brennan.2020}, a more
comprehensive and mechanistic understanding of dispersal and connectivity is
missing (but see \citealp{Creel.2020}). Nevertheless, with about 6,000
free-ranging wild dogs remaining in fragmented subpopulations
\citep{Woodroffe.2012}, reliable information on dispersal behavior and landscape
connectivity is essential for the conservation of this endangered carnivore. We
anticipated that a connectivity assessment based upon our three-step approach
would overcome several of the conceptual shortcomings of traditional
connectivity models, while providing a more detailed view on movement behavior
during dispersal its implications for landscape connectivity.

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width = \textwidth]{99_GraphicalAbstract.pdf}
    \caption{Flowchart of the simulation-based connectivity analysis. First, GPS
    data and habitat covariates must be collected. The combined data is then
    analyzed using an integrated step selection model (step 1). The parametrized
    model is then treated as an individual-based movement model and used to
    simulate dispersal trajectories (step 2). Ultimately, simulated trajectories
    are translated into a set of maps that are pertinent to landscape
    connectivity (step 3). This includes a heatmap, indicating the traversal
    frequency across each spatial unit of the study area, a betweenness map,
    highlighting movement corridors and bottlenecks, and, finally, an
    inter-patch connectivity map, where the frequency of connections and their
    average duration can be depicted.}
    \label{GraphicalAbstract}
  \end{center}
\end{figure}

\section{Methods}
\subsection{Case Study}
\subsubsection{GPS Data}
We applied the three step approach presented in \Cref{GraphicalAbstract} to GPS
movement data from 16 dispersing African wild dog coalitions (7 female and 9
male coalitions). This data has been collected between 2011 and 2019 from a
free-ranging wild dog population in northern Botswana. During dispersal, GPS
collars recorded a fix every 4 hours and regularly transmitted data over the
Iridium satellite system. To ensure comparable time intervals between GPS fixes,
we removed any fixes that were not successfully obtained at the desired 4-hour
schedule (allowing for a tolerance of \( \pm \) 15 minutes). To prepare the data
for step-selection analysis, we converted the fixes (n = 4'169) into steps,
where each step represented the straight-line movement between two consecutive
GPS fixes \citep{Turchin.1998}. We only considered steps with equal
step-durations (i.e. 4 hours) for further analysis. We will refer to these steps
as ``realized steps''. We did not differentiate between sexes, for previous
research found little differences between sexes during dispersal
\citep{Woodroffe.2019, Cozzi.2020}. Additional details on the data collection
and preparation can be found in \cite{Cozzi.2020} and \cite{Hofmann.2021}.

@@

\subsubsection{Study Area}
Our simulation of dispersal trajectories and assessment of connectivity spanned
across the entire Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA,
\Cref{StudyArea}a and b) and encompassed a rectangular extent of roughly 1.3
Mio. km\textsuperscript{2}. With an area of 520'000 km\textsuperscript{2}, the
KAZA-TFCA is the world's largest transboundary conservation area and comprises
parts of Angola, Botswana, Namibia, Zimbabwe, and Zambia, thus hosting a rich
diversity of landscapes, ranging from savannah to grassland and from dry to
moist woodland habitats. In its center lies the Okavango Delta, a dominant
hydro-geographical feature and the world's largest flood-pulsing inland delta.
Large portions of the KAZA-TFCA are formally protected in the form of national
parks (NPs) or other protected areas, yet a considerable portion of the
landscape remains human-dominated (e.g. roads, agricultural sites, and
settlements).

\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
        \includegraphics[width=\textwidth]{99_StudyArea.png}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
            % % next four lines will help you to locate the point needed by forming a grid.
            % % comment these four lines in the final picture.
            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
            % % upto here
            \draw[densely dotted, blue] (0.169, 0.222) -- (0.364, 0.955);
            \draw[densely dotted, blue] (0.169, 0.157) -- (0.364, 0.074);
        \end{scope}
    \end{tikzpicture}
    \caption{Illustration of the study area in southern Africa. (a) The study
    area was confined by a bounding box spanning the entire KAZA-TFCA which
    comprises parts of Angola, Namibia, Botswana, Zimbabwe, and Zambia. Data on
    remaining wild dog populations (orange) has been sourced from
    \cite{Woodroffe.2012}. (b) The KAZA-TFCA represents the world's largest
    terrestrial transfrontier conservation area and covers a total area of
    520'000 km\textsuperscript{2}. Its main purpose is to re-establish
    connectivity between already-existing NPs (dark green) and other protected
    areas (light green).}
    \label{StudyArea}
  \end{center}
\end{figure}

>>>>

\documentclass[../FinalThesis.tex]{subfiles}
\begin{document}
\chapter{A Three-Step Approach for Assessing Landscape Connectivity via
Simulated Dispersal: African Wild Dog Case Study}
\label{ChapterSimulation}
\thispagestyle{empty}
\vspace{-1cm}
\noindent\hfil\rule{0.75\textwidth}{.4pt}\hfil

\begin{center}
David D. Hofmann \orcid{0000-0003-3477-4365},
Gabriele Cozzi \orcid{0000-0002-1744-1940},
John W. McNutt,
Arpat Ozgul \orcid{0000-0001-7477-2642}, and
Dominik M. Behr \orcid{0000-0001-7378-8538}

% \begin{flushleft}

% \vspace{5cm}
\vfill

Published in Landscape Ecology (2023):
\url{https://doi.org/10.1007/s10980-023-01602-4}.

% \textsuperscript{1} Department of Evolutionary Biology and Environmental
% Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
% Switzerland.
%
% \vspace{0.25cm}
%
% \textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
% Private Bag 13, Maun, Botswana.
%
% \vspace{0.25cm}
%
% \textsuperscript{\S} Corresponding author:
% \href{mailto:david.hofmann2@uzh.ch}{david.hofmann2@uzh.ch}

\end{center}

% \vfill

% \textbf{Running Title:}\\African Wild Dog Dispersal in a Changing Climate:
% Lessons Learned from Seasonal Flood Extremes
%
% \vspace{0.5cm}
%
% \textbf{Keywords:}\\movement, connectivity, climate change, seasonality,
% dispersal, conservation, individual-based simulations, step-selection

% \end{flushleft}

\newpage
\section*{Abstract}
Dispersal of individuals contributes to long-term population persistence, yet
requires a sufficient degree of landscape connectivity. To date, connectivity
has mainly been investigated using least-cost analysis and circuit theory, two
methods that make assumptions that are hardly applicable to dispersal. While
these assumptions can be relaxed by explicitly simulating dispersal trajectories
across the landscape, a unified approach for such simulations is lacking. Here,
we propose and apply a simple three-step approach to simulate dispersal and to
assess connectivity using empirical GPS movement data and a set of habitat
covariates. In step one of the proposed approach, we use integrated
step-selection functions to fit a mechanistic movement model describing habitat
and movement preferences of dispersing individuals. In step two, we apply the
parameterized model to simulate dispersal across the study area. In step three,
we derive three complementary connectivity maps; a heatmap highlighting
frequently traversed areas, a betweenness map pinpointing dispersal corridors,
and a map of inter-patch connectivity indicating the presence and intensity of
functional links between habitat patches. We demonstrate the applicability of
the proposed three-step approach in a case study in which we use GPS data
collected on dispersing African wild dogs (\textit{Lycaon pictus}) inhabiting
northern Botswana. Using step-selection functions we successfully parametrized a
detailed dispersal model that described dispersing individuals' habitat and
movement preferences, as well as potential interactions among the two. The model
substantially outperformed a model that omitted such interactions and enabled us
to simulate 80,000 dispersal trajectories across the study area. By explicitly
simulating dispersal trajectories, our approach not only requires fewer
unrealistic assumptions about dispersal, but also permits the calculation of
multiple connectivity metrics that together provide a comprehensive view of
landscape connectivity. In our case study, the three derived connectivity maps
revealed several wild dog dispersal hotspots and corridors across the extent of
our study area. Each map highlighted a different aspect of landscape
connectivity, thus emphasizing their complementary nature. Overall, our case
study demonstrates that a simulation-based approach offers a simple yet powerful
alternative to traditional connectivity modeling techniques. It is therefore
useful for a variety of applications in ecological, evolutionary, and
conservation research.

\newpage
\section{Introduction}

% Importance of Dispersal & Connectivity
Dispersal of individuals is a vital process that allows species to maintain
genetic diversity \citep{Perrin.2000, Frankham.2002, Leigh.2012, Baguette.2013,
LaPoint.2013}, rescue non-viable populations \citep{Brown.1977}, and to colonize
unoccupied habitats \citep{Hanski.1999a, MacArthur.2001}. However, the ability
to disperse depends on a sufficient degree of landscape connectivity
\citep{Fahrig.2003, Clobert.2012}, making the identification and protection of
dispersal corridors that promote connectivity a task of fundamental importance
\citep{Doerr.2011, Rudnick.2012}. Identifying dispersal corridors not only
necessitates a comprehensive understanding of the factors that limit dispersal,
but also an appropriate model to estimate connectivity \citep{Baguette.2013,
Vasudev.2015, Hofmann.2021}. To date, the most commonly used connectivity models
are least-cost path analysis (LCPA; \citealp{Adriaensen.2003}) and circuit
theory (CT; \citealp{McRae.2006, McRae.2008}). Unfortunately, both models rest
on assumptions that appear unsuitable for dispersers, thus calling for the
development of alternative approaches. One promising alternative is to assess
landscape connectivity via simulated dispersal trajectories generated from
individual-based movement models (IBMMs, \citealp{Diniz.2019}). However, IBMMs
require a large number of subjective modeling decisions, thus making
among-system comparisons difficult.

% Issues with Both Methods
Traditional connectivity models make assumptions that are rarely met for
dispersers. LCPA, for instance, assumes that individuals move towards a
preconceived endpoint and choose a cost-minimizing route accordingly
\citep{Sawyer.2011, Abrahms.2017}. While this assumption may be justifiable for
migrating animals, it is unlikely to hold for dispersers, as dispersers
typically move across unfamiliar territory towards an unknown endpoint
\citep{Koen.2014, Cozzi.2020}. CT, on the contrary, posits that animals move
according to a random walk, entailing that autocorrelation between subsequent
movements cannot be rendered \citep{Diniz.2019}. For dispersers, however,
autocorrelated movements are regularly observed \citep{Cozzi.2020,
Hofmann.2021}, meaning that dispersal trajectories are usually strongly
directional. An interesting generalization that bridges the continuum between
LCPA and CT has been proposed by \citet{Panzacchi.2016} and enables to
capitalize on the merits of both approaches. Despite these and several other
generalizations of LCPA and CT (e.g. \citealp{Pinto.2009, Landguth.2012,
Panzacchi.2016, Brennan.2020}), some shortcomings remain. Most notably, all of
these methods rely on static permeability or resistance surfaces that can't
reflect the temporal dimension of dispersal. This permits statements about the
expected duration for moving between habitat patches \citep{Martensen.2017,
Diniz.2019}.

% What about IBMMs?
The shortcomings inherent to LCPA and CT can be overcome by simulating dispersal
using IBMMs and by converting simulated trajectories into meaningful measures of
connectivity \citep{Diniz.2019}. In contrast to LCPA and CT, IBMMs allow to
explicitly simulate how individuals move across and interact with the
encountered landscape \citep{Kanagaraj.2013, Clark.2015, Allen.2016,
Hauenstein.2019, Zeller.2020}, as well as to render potential interactions
between movement behavior and habitat conditions \citep{Avgar.2016}. This shifts
the focus towards a more functional view on connectivity
\citep{Tischendorf.2000}. Furthermore, IBMMs generate movement sequentially,
i.e. they generate a series of steps, so that the temporal dimension of
dispersal becomes explicit and allows modeling autocorrelation between
successive steps \citep{Diniz.2019}. Finally, simulations from IBMMs do not
enforce movement or connections towards preconceived endpoints but allow
individuals to adjust their route  ``on the go'', thereby preventing biases
arising from misplaced endpoints. Despite these advantages, a unifying approach
to simulate dispersal and assess connectivity using IBMMs is lacking.
Considering the large number of subjective decisions entailed by IBMMs, an
approach that streamlines and standardizes the application of dispersal
simulations to assess connectivity will, however, be critical to safeguard
comparability among studies.

% Proposed Solution: Three-Step approach
Here, we propose and exemplify a simple three-step approach for simulating
dispersal and assessing landscape connectivity (\Cref{GraphicalAbstract}). In
step one, we combine GPS movement data of dispersing individuals with habitat
covariates to fit a mechanistic movement model via integrated step-selection
functions (iSSFs, \citealp{Avgar.2016}). We chose to use iSSFs because the
framework not only allows inference on the study species' habitat kernel (i.e.
its habitat preferences), but also its movement kernel (i.e. its movement
preferences/capabilities) and potential interactions among the two
\citep{Avgar.2016, Fieberg.2021}. In step two, we use the parametrized movement
model to simulate dispersal across the study area. Comparable simulations have
already been applied to estimate steady-state utilization distributions of
resident individuals \citep{Potts.2013, Signer.2017} and to model landscape
connectivity, yet disregarding interdependencies between habitat and movement
kernels \citep{Clark.2015, Zeller.2020}. Finally, in step three, we convert the
simulated trajectories into three complementary connectivity maps; (i) a heatmap
revealing frequently traversed areas (e.g. \citealp{Hauenstein.2019,
Zeller.2020}), (ii) a betweenness-map delineating dispersal corridors and
bottlenecks (e.g. \citealp{Bastille-Rousseau.2018}), (iii) and a map of
inter-patch connectivity, depicting the presence and intensity of functional
links between habitat patches, as well as the average dispersal duration
required to realize those connections (e.g. \citealp{Gustafson.1996,
Kanagaraj.2013}).

@@

% Introduction of the study species
We showcase the application of the proposed approach using GPS movement data
collected on dispersing African wild dogs (\textit{Lycaon pictus}). The African
wild dog is a highly mobile species whose population persistence heavily relies
on the availability of large, natural or semi-natural landscapes and a
sufficient degree of connectivity among remaining subpopulations. Once common
throughout sub-Saharan Africa, this species has disappeared from much of its
historic range, largely due to human persecution, habitat fragmentation, and
disease outbreaks \citep{Woodroffe.2012}. Wild dogs typically disperse in
single-sex coalitions \citep{McNutt.1996, Behr.2020} and are capable of
dispersing several hundred kilometers \citep{Davies-Mostert.2012, Masenga.2016,
Cozzi.2020}. Although previous studies have investigated connectivity for this
species using LCPA \citep{Hofmann.2021} and CT \citep{Brennan.2020}, a more
comprehensive and mechanistic understanding of dispersal and connectivity is
missing (but see \citealp{Creel.2020}). Nevertheless, with about 6,000
free-ranging wild dogs remaining in fragmented subpopulations
\citep{Woodroffe.2012}, reliable information on dispersal behavior and landscape
connectivity is essential for the conservation of this endangered carnivore. We
anticipated that a connectivity assessment based upon our three-step approach
would overcome several of the conceptual shortcomings of traditional
connectivity models, while providing a more detailed view on movement behavior
during dispersal its implications for landscape connectivity.

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width = \textwidth]{Figures/GraphicalAbstract}
    \caption{Flowchart of the simulation-based connectivity analysis. First, GPS
    data and habitat covariates must be collected. The combined data is then
    analyzed using an integrated step selection model (step 1). The parametrized
    model is then treated as an individual-based movement model and used to
    simulate dispersal trajectories (step 2). Ultimately, simulated trajectories
    are translated into a set of maps that are pertinent to landscape
    connectivity (step 3). This includes a heatmap, indicating the traversal
    frequency across each spatial unit of the study area, a betweenness map,
    highlighting movement corridors and bottlenecks, and, finally, an
    inter-patch connectivity map, where the frequency of connections and their
    average duration can be depicted.}
    \label{GraphicalAbstract}
  \end{center}
\end{figure}

\section{Materials and Methods}
\subsection{Case Study}
\subsubsection{GPS Data}

We applied the three step approach presented in \Cref{GraphicalAbstract} to GPS
movement data from 16 dispersing African wild dog coalitions (7 female and 9
male coalitions). This data has been collected between 2011 and 2019 from a
free-ranging wild dog population in northern Botswana. During dispersal, GPS
collars recorded a fix every 4 hours and regularly transmitted data over the
Iridium satellite system. To ensure comparable time intervals between GPS fixes,
we removed any fixes that were not successfully obtained at the desired 4-hour
schedule (allowing for a tolerance of \( \pm \) 15 minutes). To prepare the data
for step-selection analysis, we converted the fixes (n = 4'169) into steps,
where each step represented the straight-line movement between two consecutive
GPS fixes \citep{Turchin.1998}. We only considered steps with equal
step-durations (i.e. 4 hours) for further analysis. We will refer to these steps
as ``realized steps''. We did not differentiate between sexes, for previous
research found little differences between sexes during dispersal
\citep{Woodroffe.2020, Cozzi.2020}. Additional details on the data collection
and preparation can be found in \citet{Cozzi.2020} and \citet{Hofmann.2021}.

@@

\subsubsection{Study Area}

Our simulation of dispersal trajectories and assessment of connectivity spanned
across the entire Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA,
\Cref{StudyAreaCH1}a and b) and encompassed a rectangular extent of roughly 1.3
Mio. km\textsuperscript{2}. With an area of 520,000 km\textsuperscript{2}, the
KAZA-TFCA is the world's largest transboundary conservation area and comprises
parts of Angola, Botswana, Namibia, Zimbabwe, and Zambia, thus hosting a rich
diversity of landscapes, ranging from savannah to grassland and from dry to
moist woodland habitats. In its center lies the Okavango Delta, a dominant
hydro-geographical feature and the world's largest flood-pulsing inland delta.
Large portions of the KAZA-TFCA are formally protected in the form of national
parks (NPs) or other protected areas, yet a considerable portion of the
landscape remains human-dominated (e.g. roads, agricultural sites, and
settlements).

\begin{figure}[htbp]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
        \includegraphics[width=\textwidth]{Figures/StudyArea.png}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
            % % next four lines will help you to locate the point needed by forming a grid.
            % % comment these four lines in the final picture.
            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
            % % upto here
            \draw[densely dotted, blue] (0.169, 0.222) -- (0.364, 0.955);
            \draw[densely dotted, blue] (0.169, 0.157) -- (0.364, 0.074);
        \end{scope}
    \end{tikzpicture}
    \caption{Illustration of the study area in southern Africa. (a) The study
    area was confined by a bounding box spanning the entire KAZA-TFCA which
    comprises parts of Angola, Namibia, Botswana, Zimbabwe, and Zambia. Data on
    remaining wild dog populations (orange) has been sourced from
    \citet{Woodroffe.2012}. (b) The KAZA-TFCA represents the world's largest
    terrestrial transfrontier conservation area and covers a total area of
    520,000 km\textsuperscript{2}. Its main purpose is to re-establish
    connectivity between already-existing NPs (dark green) and other protected
    areas (light green).}
    \label{StudyAreaCH1}
  \end{center}
\end{figure}

\subsubsection{Habitat Covariates}
